Critical assessment of Linear-Response Excitations in Molecular system\\ or \\ On the Localization of the Linear-Response excitations in Open Systems

### 2018-01-19 LRDraft.tex
update

\subsection{Localization argument for the fluctuation states}

@@

We split the unperturbed Hamiltonian as the kinetic $\op T$ plus the potential $\op V$ term, so that the equation of motion for $\ket{f_p(\omega)}$ is recasted as:
\be\lb{fluctuationStateEqMotion}
\left(\bbomega+\eps_p -\op T\right)\ket{f_p(\omega)} = \op V\ket{f_p(\omega)} + \op Q_0\left(\op V'(\omega) + \op \Phi(\omega)\right)\ket{\psi_p} \;.
\ee
This equation has the formal structure of an inhomogeneous Helmholtz equation and can be written in the coordinate representation as:
\be
\left(k_p(\omega)^2+\ii\eta + \nabla^2\right)f_p(\omega,\r) = s(\omega,\r) \;\;, {\rm with} \;\; k_p(\omega)^2 = 2(\omega+\eps_p) \;,
\ee
and admits an integral representation in the form
\be\lb{fluctuationStateIntEq1}
f_p(\omega,\r) = \int \dd\r' g_H(\r-\r')s(\r') \;,
\ee
where $g_H(\r)$ is the Green function of the Helmholtz operator in three dimensions. This integral representation allows us to assess the localization character 
of the fluctuation states, \emph{i.e.} the behavior of $f_p(\omega,\r)$ when $\r$ becomes large. Indeed, assuming that the support of the source term is compact
we have that, for $\r$ outside the domain in which the integrand of \eqref{fluctuationStateIntEq1} is not vanishing:
\be\lb{fluctuationStateIntEq2}
f_p(\omega,\r) \sim \frac{e^{\ii \sqrt{k_p(\omega)^2} |\r|}}{|\r|} \;,
\ee
so the asymptotic character of the fluctuation states depends on the sign of $k_p(\omega)^2$. In particular equation \eqref{fluctuationStateIntEq2} describes an
exponential damping for $\omega < |\eps_p|$ and an oscillating power law behavior for $\omega \geq |\eps_p|$.

@@

%The linear response presented so far is expressed in functions of the fluctuation states. These objects essentially describe the modification of the occupied eigenstates of 
%$\hnot$ induced by the perturbation. They exhibit an explicit dependence on the perturbing field and their $\omega$ dependence is a complicated combination of the intrinsic
%feature of the system

The linear response presented so far is expressed in functions of the fluctuation states. These objects essentially describe the modification of the occupied eigenstates of 
$\hnot$ induced by the perturbation and, by definition, explicitly depend on the perturbing field. As a consequence, the $\omega$ dependence of fluctuation states is a complicated 
combination of the one of $\op\Phi(\omega)$ and of the intrinsic energetic feature of the system.

For these reason it is preferable to present an equivalent but alternative formula for the linear response functional \eqref{LinearResponseFunctDef2}...

%They exhibit an explicit dependence on the perturbing field and their $\omega$ dependence is a complicated combination of the intrinsic
%eature of the system

>>>>

\subsection{Localization argument for the fluctuation states}
\lb{LocalizationFluctuationStates}

@@

We split the unperturbed Hamiltonian as the kinetic $\op T$ plus the potential $\op V$ term, so that the equation of motion for $\ket{f_p(\omega)}$ is recasted as:
\be\lb{fluctuationStateEqMotion}
\left(\bbomega+\eps_p -\op T\right)\ket{f_p(\omega)} = \op V\ket{f_p(\omega)} + \op Q_0\left(\op V'(\omega) + \op \Phi(\omega)\right)\ket{\psi_p} \;.
\ee
This equation has the formal structure of an inhomogeneous Helmholtz equation and can be written in the coordinate representation as:
\be
\left(k^2_p(\omega)+\ii\eta + \nabla^2\right)f_p(\omega,\r) = s(\omega,\r) \;\;, {\rm with} \;\; k^2_p(\omega) = 2(\omega+\eps_p) \;,
\ee
and admits an integral representation in the form
\be\lb{fluctuationStateIntEq1}
f_p(\omega,\r) = \int \dd\r' g_H(\r-\r')s(\r') \;,
\ee
where $g_H(\r)$ is the Green function of the Helmholtz operator in three dimensions. This integral representation allows us to assess the localization character 
of the fluctuation states, \emph{i.e.} the behavior of $f_p(\omega,\r)$ when $\r$ becomes large. Indeed, assuming that the support of the source term is compact
we have that, for $\r$ outside the domain in which the integrand of \eqref{fluctuationStateIntEq1} is not vanishing:
\be\lb{fluctuationStateIntEq2}
f_p(\omega,\r) \sim \frac{e^{\ii \sqrt{k^2_p(\omega)} |\r|}}{|\r|} \;,
\ee
so the asymptotic character of the fluctuation states depends on the sign of $k^2_p(\omega)$. In particular equation \eqref{fluctuationStateIntEq2} describes an
exponential damping for $\omega < |\eps_p|$ and an oscillating power law behavior for $\omega \geq |\eps_p|$.

@@

The linear response presented so far is expressed in functions of the fluctuation states, that essentially describe the modification of the occupied eigenstates of $\hnot$ induced by 
the perturbation. Fluctuation states depend explicitly on the perturbing field $\op\Phi(\omega)$ and, consequently, the analytic structure of the linear response written as in formula 
\eqref{LinearResponseFunctDef2} emerge through a combination of the $\omega$ dependence of the perturbing field an on the intrinsic response of the system under inspection.

It is however possible to present and equivalent but alternative formulation of the linear response that allows us to express the response density operator in term of the spectral
decomposition of the Liouville superoperator. This approach naturally leads to the concept of \emph{excitation operators}, defined through a self-consistent operator eigenvalue equation,
that codify the intrinsic capability of the system to respond to an arbitrary perturbation. Excitations behave as singularity of the response density operator and their knowledge provides a 
deeper insight into the analytic structure of the linear response.

The material presented in this section discusses a complete re-derivation of the linear response written in terms of excitation operators. Moreover, we will evidence that a \emph{localization 
argument}, analogous to the one presented in section \ref{LocalizationFluctuationStates} for the fluctuation states, holds also for the excited states of the systems below a specific energetic
threshold. This analysis has relevant consequence both under theoretical and computational point of view. Indeed, it provides a deeper understanding in the role of both localized and de-localized 
excitations in determining the analytic structure of the linear response and it gives us clear indication in the level of refinement of the computational setup needed to obtain reliable results 
from a linear response calculation.
